---
title: "Python-to-R Data Pipeline Demonstration"
author: "Alec Brooks"
date: today

format:
  html:
    theme: darkly
    toc: true
    code-fold: false

execute:
  warning: false
  message: false
---

# Overview

This report demonstrates a reproducible mixed-language data pipeline using Python, R, and Quarto.

Python is used to retrieve and prepare the source data, including selecting relevant variables and removing incomplete observations. The cleaned pandas DataFrame is then transferred to R for statistical modeling, inference, visualization, and prediction. Quarto combines each stage into a single reproducible report.

# Setup

## R: Python integration

This is required for Python to pass off data to R within a Quarto report.

```{r}
#| label: configure-python

library(reticulate)
use_python("/usr/bin/python", required = TRUE)
```

## Python: Data acquisition

The Palmer Penguins dataset is downloaded directly from its public repository, allowing the report to be reproduced without relying on a local database or file.

```{python}
#| label: download-data

import pandas as pd

url = "https://raw.githubusercontent.com/allisonhorst/palmerpenguins/main/inst/extdata/penguins.csv"

penguins = pd.read_csv(url)

penguins.head()
```

## Python: Data preparation

Only the variables required for the analysis are retained. Observations containing missing values in these variables are removed.

```{python}
#| label: clean-data

penguins_clean = (
    penguins[
        [
            "species",
            "flipper_length_mm",
            "body_mass_g"
        ]
    ]
    .dropna()
)

penguins_clean.head()
```

# Analysis

## R: Data transfer

The cleaned pandas DataFrame is transferred from the Python session into R using `reticulate`.

```{r}
#| label: transfer-data

penguins_r <- py$penguins_clean
```

## R: Linear regression

A multiple linear regression model is fitted to estimate body mass using flipper length and species.

```{r}
#| label: fit-model

model <- lm(
  body_mass_g ~ flipper_length_mm + species,
  data = penguins_r
)

```

The coefficient p-values are extracted separately for concise presentation.

```{r}
#| label: model-p-values

summary(model)$coefficients[, "Pr(>|t|)"]
```

# Visualization

The scatter plot shows the relationship between flipper length and body mass for each penguin species. Separate regression lines are fitted within each species.

```{r}
#| label: fig-flipper-mass-species
#| fig-cap: "Relationship between flipper length and body mass by species"

library(ggplot2)

ggplot(
  penguins_r,
  aes(
    x = flipper_length_mm,
    y = body_mass_g,
    color = species
  )
) +
  geom_point(alpha = 0.7, size = 2) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Penguin Body Mass by Flipper Length",
    x = "Flipper length (mm)",
    y = "Body mass (g)",
    color = "Species"
  ) +
  theme_minimal()
```

# Prediction

The fitted model is used to predict body mass for new combinations of flipper length and species. Prediction intervals represent the expected range for individual penguins with these characteristics.

```{r}
#| label: predictions

new_penguins <- expand.grid(
  flipper_length_mm = c(180, 200, 220),
  species = c("Adelie", "Chinstrap", "Gentoo")
)

predictions <- predict(
  model,
  newdata = new_penguins,
  interval = "prediction"
)

prediction_table <- cbind(
  new_penguins,
  predictions
)

prediction_table
```

# Conclusion

This report demonstrated a reproducible mixed-language workflow in which Python was used to retrieve and clean the Palmer Penguins dataset before passing the prepared data to R for analysis.

The regression results showed that flipper length is a strong predictor of body mass. Species also contributed additional explanatory value, reflecting meaningful differences in body size among Adelie, Chinstrap, and Gentoo penguins. The visualization reinforced these findings by showing a clear positive relationship between flipper length and body mass within each species.

The fitted model was then used to generate body-mass predictions for new observations. Together, these steps demonstrate how Python, R, and Quarto can be combined into a single reproducible pipeline for data acquisition, preparation, statistical modeling, visualization, and prediction.